---
title: "Inference"
format: 
  revealjs:
    incremental: true
    embed-resources: true
    slide-number: true
    html-math-method: mathjax
editor: visual
execute: 
  echo: true
---

## Announcements

-   HW 2 due Sunday

-   Quiz 1 on Tuesday: covers all of unit 1, no code

## What is hypothesis testing?

Hypothesis testing is the procedure that assesses the evidence from the data in favor of or against some **claims (hypotheses) about the population**.

 

We will use information from the **sample (data)** to draw conclusions about the **population**

## What is hypothesis testing?

Usually we have two hypotheses to test:

-   Null ($H_0$): Formally, the null hypothesis makes a **claim** or **assumption** about a population parameter. Conceptually, it's the "nothing special is going on" hypothesis.

-   Alternative ($H_A$): A statement that is contradictory to the null hypothesis. Conceptually, it is the "something special is going on" or "the variables are associated" hypothesis.

## Poll

As we test the hypotheses, do you think it would be better scientific practice to operate under the assumption that the null hypothesis is true (i.e., nothing special is going on) or that the alternative hypothesis is true (something special is going on/the variables are related)?

For example, imagine you work for the FDA and you are assessing whether or not a new drug should be approved. Do you think it is better practice to 1) assume that the drug **is not** effective until the data prove otherwise, or 2) assume that the drug **is** effective until the data prove otherwise?

## Hypothesis testing framework

-   Start with two hypotheses about the population: $H_0$ (null) and $H_A$ (alternative).
-   Choose a representative sample.
-   Assess **how likely** the sample data can be observed **given that the** $H_0$ is true.
-   If the data are very unlikely to be observed when $H_0$ is true, then we reject $H_0$; otherwise, we cannot reject the null hypothesis.

## P-value

The probability of **observing the data or even more extreme values** assuming that $H_0$ is true.

-   $H_0:\mu=0;H_A:\mu>0$.
-   Then p-value $=P(X>x_{obs})$ when $\mu=0$
-   It is **NOT** the probability of $H_0$ is true
-   What if $H_A: \mu \neq 0$ ?

## Example

You want to determine if a coin is a biased coin. You perform an experiment with the coin in question.

1.  Specify the null and alternative hypotheses
2.  Collect data
3.  Calculate the p-value

## Example

Data: TTHTHHTHHH

 

. . .

Calculate p-value:

```{r}
2*(1-pbinom(6,10,0.5))
```

Do you think the coin is biased?

. . .

What if we increased our sample size but saw the same proportion?

```{r}
2*(1-pbinom(60,100,0.5))
```

Do you think the coin is biased?

## Significance level

A predetermined threshold used to tell whether a result is statistically significant; usually denoted by $\alpha$ and set as 0.05.

-   If the p-value is smaller than $\alpha=0.05$, then we can reject $H_0$ at 0.05-significant level.
-   If the p-value is greater than $\alpha=0.05$, then we fail to reject $H_0$ at 0.05-significant level.

. . .

Note: **We can NEVER say we accept the null hypothesis**, even if the p-value is greater than the significance level.

## Null Distribution Illustration

## Simulation-based Inference vs. Parametric Inference

Simulation-based inference

-   Simulate the null distribution with resampling methods

-   Very flexible, but requires computation power

. . .

Parametric Inference

-   Make assumptions about the probability distribution of our population

-   Relies on statistical theory such as central limit theorem

## Why learn both approaches?

-   Simulation-based inference can be intuitive, and is more flexible

-   Parametric inference has been around for a long time and you will likely see it more in practice

. . .

Note: There is also **non-parametric inference**, where we don't make distributional assumptions (e.g., Wilcoxon tests)

## An ethical decision
