---
title: "IDS 702 HW 5"
author: "Your Name Here"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems; it is insufficient to just show the code.

**Add your name in the Author section in the header**

## Load data

```{r}
library(MASS)
library(tidyverse)
library(glmmTMB)
## NOTE: Run install.packages("glmmTMB", type = "source") IN CONSOLE

disaster_deaths <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/disaster_deaths_cleaned.csv")
```

## 1

a\.

```{r}

```

b\.

```{r}

```

## 2

a\.

```{r}

```

b\.

```{r}

```

## 3

```{r}

```

## 4

```{r}
disaster_deaths <- disaster_deaths |>
  mutate(Year_1 = Year - min(Year)+1)

zip_model_pois <- glmmTMB(
  Deaths ~ Region*Year_1,
  ziformula = ~ Region*Year_1,
  family = poisson,
  data = disaster_deaths
)

summary(zip_model_pois)
```

a\.

```{r}

```

b\.

```{r}

```

### Bonus
