---
title: "Inference"
subtitle: "IDS 702"
format: html
editor: visual
---

## Load packages

```{r load-libraries, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

### Data

```{r load-data, message = F, warning = F}
push_pull <- read_csv("https://sta101-fa22.netlify.app/static/appex/data/push_pull.csv")
```

```{r peak-at-data}
push_pull |>
  slice(1:3, 24:26)
```

The `push_pull` dataset comes from a "mini study" by [mountain tactical institute](https://mtntactical.com/knowledge/mini-study-grease-the-groove-beats-density-for-push-up-pull-up-improvement/).

26 individuals completed 1 of 2 exercise regiments for 3.5 weeks to increase their pushups and pullups. Codebook below:

-   `participant_id`: unique identifier for each participant
-   `age`: age of participant
-   `push1/2`: push-ups at beginning and end of program respectively
-   `pull1/2`: pull-ups at beginning and end of program respectively
-   `training`: which training protocol the individual participated in

## Summary

### Hypothesis testing procedure

1.  Specify the null and alternative hypothesis. Choose or know $\alpha$.
2.  Collect/examine the data. Compute the observed statistic.
3.  Simulate **under the null** and compute the p-value using the **observed statistic** and the **alternative hypothesis**.
4.  Compare the p-value to your significance level $\alpha$ and reject or fail to reject the null. Interpret your result in context.

We typically also compute a confidence interval using the observed statistic for additional information about the parameter of interest.

## Hypothesis test: Which training method is better?

Two exercise regimes:

-   "density" training
-   "grease-the-groove" (gtg)

We want to know, is the average pull-up percent increase of a gtg trainee significantly different than a density trainee?

In other words, does the **categorical variable** `training` affect the **numeric variable** average percentage increase in pull-ups?

**State the null hypothesis**:

$H_0: \mu_d = \mu_{gtg}$

In other words: $H_0: \mu_d-\mu_{gtg}=0$

$H_A: \mu_d-\mu_{gtg} \neq 0$

**Assumption:** The observations (within and across groups!) are independent

First, we'll compute the observed statistic.

```{r, results = 'hide'}
push_pull <- push_pull |>
  mutate(pct_pull_inc = (pull2 / pull1) - 1)

# observed statistic
obs_stat <- push_pull |>
  specify(response = pct_pull_inc, #response=numeric
          explanatory = training) |> #exp=categorical
  calculate(stat = "diff in means", 
            order = c("density", "gtg"))
```

Now we'll simulate under the null and compute the p-value:

```{r}
set.seed(1)
# simulate null
# null distribution here is the distribution of the difference in the two means when there is no relationship between the categorical and numeric variables
# "permute" shuffles the training assignments to "break" the potential relationship between training and pct increase
null_dist <- push_pull |>
  specify(response = pct_pull_inc, 
          explanatory = training) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", 
            order = c("density", "gtg"))

# get p-value
p_value <- null_dist |>
  get_p_value(obs_stat, direction = "both")

# visualize
visualize(null_dist) +
  shade_p_value(obs_stat, direction = "both")
```

-   State your conclusion with $\alpha$= 0.05

-   Generate a bootstrap confidence interval for the mean difference of density and gtg training percent increase in pull ups

```{r, eval=FALSE}
set.seed(1)
boot_dist <- push_pull |>
  specify(response = pct_pull_inc, 
          explanatory = training) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means", 
            order=c("density", "gtg"))

get_confidence_interval(boot_dist, level=0.95, 
                        type="percentile")
```

## Hypothesis test: Is the median pull-up percent increase of a gtg trainee significantly greater than a density trainee?

$H_0: \eta_d=\eta_{gtg}$ or $H_0: \eta_{gtg}-\eta_{d} = 0$

$H_A: \eta_{gtg}>\eta_{d}$ or $\eta_{gtg}-\eta_{d}>0$

```{r, eval=FALSE}
obs_stat <- push_pull |>
  specify(response = pct_pull_inc, #response=numeric
          explanatory = training) |> #exp=categorical
  calculate(stat = "diff in medians", 
            order = c())

#simulate null distribution
set.seed(1)
null_dist <- push_pull |>
  specify(response = pct_pull_inc, 
          explanatory = training) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in medians", 
            order = c())

# get p-value
p_value <- null_dist |>
  get_p_value(obs_stat, direction = "")
```
