---
title: "IDS 702 HW 3 Part 2"
author: "Your Name Here"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems. For example, you should write "There are XX observations and each observation represents a \_\_\_\_\_"; it is insufficient to just show the code.

**Add your name in the Author section in the header**

## Load data

```{r}
library(tidyverse)

colleges <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/colleges.csv")
```

```{r}
#Code from part 1 to remove relevant missing values and factor categorical variables


```

## 1

a\.

```{r}

```

b\.

```{r}

```

c\.

```{r}

```

d\.

```{r}

```

e\.

```{r}


```

f\.

```{r}

```

g\.

h\.

## 

## 2

a\.

```{r}

```

b\.

```{r}

```

c\.

```{r}

```

d\.

```{r}

```

## Bonus
