---
title: "Confidence Intervals"
subtitle: "IDS 702"
format: html
---

## Load packages

```{r load-libraries, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

## Load Data

```{r load-data, warning = F, message = F}
manhattan <- read_csv(
  "https://sta101-fa22.netlify.app/static/appex/data/manhattan.csv"
  )
```

## Calculating confidence intervals with bootstrap

Fill in the following code to generate 1000 bootstrap means for rent price

```{r, eval=FALSE}
sample_mean <- manhattan |>
  summarise(mean(rent)) |>
  pull()

set.seed(2)
boot_dist_rent <-  |>
  specify(response = ) |>
  generate(reps = , type = ) |>
  calculate(stat= )
```

Generate the histogram of bootstrap means:

```{r, eval=FALSE}
ggplot(boot_dist_rent, aes(x=stat))+
  geom_histogram()
```

 

 

We will use this histogram/distribution of the bootstrap means to construct a confidence interval for the population mean.

Let's set our confidence level to 95% (90, 95, and 99 are standard). We want to identify the values on the x-axis of this distribution that capture 95% of the bootstrap means (i.e., 950 out of the 1000 bootstrap means).

The **quantile** function will give us these values based on a probability. For 95% confidence, we want probabilities of 0.025 and 0.975. **Why?**

```{r confidence-interval-95, eval=FALSE}
boot_dist_rent |>
  get_confidence_interval(level = 0.95, type = "percentile")
```

#### Visualize the interval

```{r, echo=FALSE, message=FALSE}
boot_CI <- boot_dist_rent |>
  get_confidence_interval(level = 0.95, type = "percentile")

ggplot(boot_dist_rent, aes(x=stat))+
  geom_histogram()+
  geom_vline(xintercept=c(boot_CI$lower_ci, boot_CI$upper_ci),
             color="red")+
  geom_vline(xintercept = sample_mean, color="blue")
```

#### Interpret the interval

If we calculated a confidence interval for each bootstrap sample, we would expect 95% of the intervals to contain the true population mean. Therefore, we can be confident that the true mean rent is between

## Confidence intervals with the central limit theorem

In general, the confidence interval can be written as

$$
\bar{x} \pm z^* \times \sigma/\sqrt{n}
$$ where $z^*$ is the quantile of a standard normal distribution associated with our level of confidence.

What about when we don't know $\sigma$ (which is nearly always the case)?

We approximate $\mu$ by $\bar{x}$ and $\sigma$ by the standard deviation $s$. However, $s$ may be smaller than $\sigma$ and our confidence interval could be too narrow. **What is the problem with a confidence interval being too narrow?**

To account for this uncertainty, we will use a distribution with thicker tails. This sampling distribution is called a **t-distribution**.

The t-distribution has a bell shape but the extra thick tails help us correct for the variability introduced by using *s* instead of *σ*.

The t-distribution, like the standard normal, is always centered at zero. Therefore, the t-distribution has only a single parameter: degrees of freedom. The degrees of freedom describes the precise form of the bell-shaped t-distribution. In general, we'll use a t-distribution with $df=n−1$ to model the sample mean when the sample size is $n$.

We can use `qt` and `pt` to find quantiles and probabilities respectively under the t-distribution.

To construct our practical confidence interval (where we don't know $\sigma$) we use the t-distribution:

$$
\bar{x} \pm t^*_{n-1} \times \frac{s}{\sqrt{n}}
$$

#### 

Calculate the 95% confidence interval for rent price using the t-distribution.

```{r clt-ci}
sample_size <- nrow(manhattan)
t.star <- abs(qt(.025,sample_size-1))
std.error <- manhattan |>
  summarise(sd(rent)/sqrt(sample_size)) |>
  pull()

sample_mean - t.star*std.error
sample_mean + t.star*std.error

```

#### Interpret the interval

If we collected many samples from the population and calculated a confidence interval for each sample, we would expect 95% of the intervals to contain the true mean rent. Therefore, we can be confident that the true mean rent is between .

How does this compare to the 95% bootstrap confidence interval?
