---
title: "IDS 702 HW 3 Part 1"
author: "Your Name Here"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems. For example, you should write "There are XX observations and each observation represents a \_\_\_\_\_"; it is insufficient to just show the code.

**Add your name in the Author section in the header**

## Load data

```{r}
library(tidyverse)

colleges <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/colleges.csv")
```

## 1

a\.

```{r}

```

b\.

```{r}

```

```{r}
colleges_full <- colleges |>
  filter(complete.cases(control, basic, student_count,
                        ft_pct, med_sat_value, endow_value,
                        grad_100_value, grad_150_value,
                        retain_value))
```

## 2

```{r}

```

|   | Count (n) | Proportion or % |
|------------------------------------------|---------------|---------------|
| **Basic classification** |  |  |
|      Baccalaureate Colleges--Arts & Sciences |  |  |
|      Baccalaureate Colleges--Diverse Fields |  |  |
|      Research Universities--high research activity |  |  |
|      Research Universities--very high research activity |  |  |
| **Control of institution** |  |  |
|      Private not-for-profit |  |  |
|      Public |  |  |

## 3

a\.

```{r}
ggplot(__________, aes(x=__________))+
  geom_histogram()+
  labs(x="Graduation rate within 100% of normal time")
```

b\.

```{r}


```

## 4

```{r, fig.height=4, fig.width=8}

ggplot(__________, aes(x=_________, y=grad_150_value,
                                   col=_________))+
  geom_point()+
  labs(x="Median SAT score",y="Grad. rate within 150% normal time",
       col="Basic classification")
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

```{r, fig.height=4, fig.width=8}

ggplot()
```

Description:

### 5

### 6

### 7
