---
title: "IDS 702 HW 3"
author: "KEY"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems. For example, you should write "There are XX observations and each observation represents a \_\_\_\_\_"; it is insufficient to just show the code.

**Add your name in the Author section in the header**

### Load Data

```{r}
library(tidyverse)
library(gridExtra)

colleges <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/colleges.csv")
```

## Part 1

### 1

a\. 797 observations and each observation is a college/university. (No code necessarily required, and can be something like `glimpse()` instead of `nrow()`)

```{r}
nrow(colleges)
```

b\.

```{r}
summary(colleges[,c("control","basic","student_count","ft_pct",
                    "med_sat_value","endow_value",
                    "grad_100_value","grad_150_value",
                    "retain_value")])
```

The following variables have missing values: