---
title: "Multinomial (Logistic) Regression"
output: html
editor: visual
---

## Generalized Linear Models

-   Allow us to fit models with different types of outcomes: binary, nominal, ordinal, count

-   Allow us to **perform inference**, i.e., obtain interpretable estimates, quantify uncertainty of those estimates, conduct hypothesis tests

## Multinomial Regression

Define $\pi_j$ as $P(Y=j)$ for a level $j$ of a nominal categorical variable.

The multinomial model is defined as a set of logistic regression models for each probability $\pi_j$ compared to a baseline level:

$$
log(\frac{\pi_{ij}}{\pi_{i1}})=\beta_{0j}+\beta_{1j}x_{i1}+...+\beta_{pj}x_{ip}
$$

What are all of the indices here ( $j$, $p$, and $i$ )?

## Interpretations

Each coefficient estimate has to be interpreted relative to the baseline outcome level

-   If $x_1$ is continuous, $e^{\beta_{1j}}$ is the multiplicative increase/decrease in the **odds** of $Y=j$ vs $Y=1$ (baseline) when increasing $x_1$ by one unit

-   If $x_1$ is categorical, $e^{\beta_{1j}}$ is the odds of $Y=j$ vs $Y=1$ (baseline) for the group with $x_1=1$ compared to the reference group

## Example

```{r, message=FALSE}
library(tidyverse)
library(caret)
library(foreign)
library(nnet)


ml <- read.dta("https://stats.idre.ucla.edu/stat/data/hsbdemo.dta")

ml <- ml |>
  mutate(prog_ac = relevel(prog, ref="academic"))

ml_mod <- multinom(prog_ac ~ ses + write, data=ml)

summary(ml_mod)

exp(summary(ml_mod)$coefficients)
```

-   The odds of being in the general track compared to academic are 41% lower for students of middle SES status compared to low.

-   A one point increase in writing score is associated with a 6% decrease in the odds of being in the general program compared to the academic program

```{r}
z <- summary(ml_mod)$coefficients/summary(ml_mod)$standard.errors

(p <- (1-pnorm(abs(z)))*2)
```

## Confusion Matrix

```{r}
confusionMatrix(predict(ml_mod), ml$prog_ac, mode="everything")
```

## 
