---
title: "Poisson Regression"
format: pdf
editor: visual
---

### The Poisson Model

We use Poisson Regression to model count outcomes. The model uses the log link function:

$$
log(\lambda_i)=\beta_{0}+\beta_1x_{i1}+...+\beta_{p}x_{ip}
$$

where $i=1,...,n$

What is $\lambda$?

### Assumptions

-   Independent observations

-   Linear relationship between $log(\lambda_i)$ and the predictors

-   **Equidispersion**

### Interpretations

-   If $x_1$ is continuous, $e^{\hat{\beta}_{1}}$ is the multiplicative increase/decrease in the **expected count of the outcome** when increasing $x_1$ by one unit

-   If $x_1$ is categorical, $e^{\hat{\beta}_{1}}$ is the multiplicative increase/decrease in the **expected count of the outcome** for $x_1=1$ compared to the reference group

## Example

Let's look at the Bikeshare data in the ISLR2 package:

```{r, message=FALSE}
library(MASS)
library(car)
library(tidyverse)
library(ISLR2)

data("Bikeshare")
# type ?Bikeshare in the console to see the data dictionary and description
```

Let's create a plot to visualize the relationship between `temp`, `season`, and `bikers`

```{r, eval=FALSE}
ggplot(Bikeshare, aes(x=temp, y=bikers, col=season))+
  geom_point()
#what's the issue here?

ggplot(Bikeshare, aes(x=temp, y=bikers, col=season))+
  geom_jitter()+
  geom_smooth(method = "glm", method.args = list(family = "poisson"))
```

### Fitting the model

```{r}
pois_mod <- glm(bikers ~ , 
                data=Bikeshare, 
                family="poisson")
summary(pois_mod)
```

```{r}
(exp_coefs <- exp(cbind(est=coef(pois_mod),confint(pois_mod))))
```

-   The expected number of bike rentals is 1.03 times higher in the spring compared to winter ( $p<0.01$, 95% CI: \[1.02,1.04\])

-   For each unit increase in normalized temperature, the expected number of bike rentals increases 15.5 times ( $p<0.01$, 95% CI: \[15.3, 15.8\])

### Assessment

Partial residual plot:

```{r}
car::crPlot(pois_mod, "temp")
```

Can compare to negative binomial model to assess equidispersion:

```{r}
nb_mod <- MASS::glm.nb(bikers ~ temp+factor(season),data=Bikeshare)

AIC(pois_mod, nb_mod)
```

## Exercise

1.  Compare the results we obtained above to a linear model:

-   Fit a linear model regressing the number of bikers on normalized temperature and season. Show the summary table and write an interpretation for the temperature predictor.

-   Generate the diagnostic plots for the linear model. Comment on what you observe, particularly in the Residuals vs. Fitted plot. Which assumption(s) appears to be violated? How would a Poisson model address this/these assumption(s)?

2.  Fit a poisson model regressing bikers on temperature and season and include an interaction term for temperature and season. Use the summary output to calculate the coefficient estimates for temperature **for each season**. Then, exponentiate those calculated estimates. Write interpretations for these values.
