---
title: "Logistic Regression Interpretation - TITANIC"
format: pdf
editor: visual
---

Answer the questions below and submit your answers on Canvas.

This exercises uses a dataset with all Titanic passengers and crew (except for the passengers distributed in class).

```{r, message=FALSE}
library(tidyverse)

titanic <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/titanic.csv")
```

1.  Provide your passenger details

    | Name | Class | Age | Sex |
    |------|-------|-----|-----|
    |      |       |     |     |

2.  Provide summary statistics (N and %) for the outcome variable (`survived`)

    ```{r}

    ```

3.  Calculate summary statistics for those with and without heart disease for the following variables: age, gender, class. (which summary statistics are appropriate for each variable?). Combine "deck crew," "engineering crew," "restaurant staff," and "victualling crew" into one class category called "crew."

    ```{r}

    ```

4.  Let's create some visuals to explore the data!

    ```{r}

    ```

5.  Now let's fit a logistic regression model regressing survival status on age, gender, and class.

    ```{r}
    titanic_mod1 <- glm(__________ ~ _________________________,
                      data=titanic,
                      family="binomial")

    summary(titanic_mod1) #log-odds scale

    exp(titanic_mod1$coefficients) #odds scale
    ```

-   What are the reference levels for the categorical variables?

-   Write interpretations for the coefficient estimates. Interpret the coefficient estimates in terms of log odds and odds/odds ratios. Are any of these results surprising?

-   Obtain confidence intervals for the coefficient estimates on the log odds and the odds scale.

    ```{r}

    ```

5.  Finally, predict your passenger's probability of survival. Then, look them up to see if the true outcome fit the probability!

    ```{r}
    passenger_details <- data.frame(age=__, gender="_____",
                                    class_combined="_____")

    predict(titanic_mod1, passenger_details, type="response")
    ```

6.  What are some interaction terms that would be interesting to add to this model?
