---
title: "HW 1"
author: "YOUR NAME HERE"
format: pdf
editor: visual
---

## Instructions

First, fill in your name in the author line in the header above. Then, use the template below to complete the assignment. For questions 1-3, you can choose to upload an image of your work rather than typing out the math (if you do want to type it, [here is a helpful resource to get started](https://qmd4sci.njtierney.com/math)). To upload an image, click "Insert \> Figure/Image."

For fraction calculations, you can simply type the fraction in a code chunk and R will show the answer when you render your document, e.g.,

```{r}
3/5
```

For problems 4 and 5, simply type the R code to complete the calculation, and the answer will appear below the code in your rendered document.

Once you have completed the exercises, render your document and submit the pdf on Gradescope. Be sure to assign the appropriate pages to the exercise numbers when you submit.

## Exercise answers

### 1

a\.

|                | Developed lung cancer | Did not develop lung cancer | Total |
|----------------|-----------------------|-----------------------------|-------|
| **Smoker**     |                       |                             |       |
| **Non-smoker** |                       |                             |       |
| **Total**      |                       |                             |       |

b\.

```{r}

```

c\.

```{r}

```

d\.

### 2

a\.

b\.

### 3

a\.

b\.

### 4

a\.

```{r}

```

b\.

```{r}

```

### 5

a\.

```{r}

```

b\.

```{r}

```

c\.

```{r}

```
