---
title: "IDS 702 HW 4"
author: "Your Name Here"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems; it is insufficient to just show the code.

**Add your name in the Author section in the header**

## Load data

```{r}
library(tidyverse)
library(openintro)
library(caret)
library(pROC)

data("resume")
```

## 1

a\.

```{r}

```

b\.

```{r}

```

+---------------------------------+-----------+-------------------+--------------------------+
| Variable                        | Overall   | Received Callback | Did not receive callback |
|                                 |           |                   |                          |
|                                 | N         | N (%)             | N(%)                     |
+=================================+===========+===================+==========================+
| **Race**                        |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Black - N (%)                   | 2435 (50) | 157 (40)          |                          |
|                                 |           |                   |                          |
| White - N (%)                   | 2435 (50) | 235 (60)          |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| **Gender**                      |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Female - N (%)                  |           |                   |                          |
|                                 |           |                   |                          |
| Male - N (%)                    |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Has college degree - N (%)      |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| **Resume Quality**              |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Low - N (%)                     |           |                   |                          |
|                                 |           |                   |                          |
| High - N (%)                    |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Years of experience - mean (SD) |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+

## 2

a\.

```{r}

```

b\.

```{r}

```

c\.

```{r}

```

## 3

a\.

```{r}

```

b\.

```{r}

```

c\.

```{r}

```

## 4
