---
title: "IDS 702 HW 4"
author: "KEY"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems; it is insufficient to just show the code.

**Add your name in the Author section in the header**

## Load data

```{r}
library(tidyverse)
library(openintro)
library(caret)
library(pROC)

data("resume")
```

## 1

a\.

```{r}
resume <- resume |>
  mutate(
    received_callback_fac = factor(received_callback,
                                   levels=c(0,1),
                                   labels=c("No","Yes")),
    race_fac = factor(race),
    gender_fac = factor(gender),
    college_degree_fac = factor(college_degree,
                                levels=c(0,1),
                                labels=c("No","Yes")),
    resume_quality_fac = factor(resume_quality)
  )

resume |>
  count(received_callback, received_callback_fac)

resume |>
  count(race, race_fac)

resume |>
  count(gender, gender_fac)

resume |>
  count(college_degree, college_degree_fac)

resume |>
  count(resume_quality, resume_quality_fac)
```

b\.

```{r}
resume |> 
  count(race_fac) |> 
  mutate(prop=n/sum(n))

resume |> 
  group_by(received_callback_fac) |> 
  count(race_fac) |> 
  mutate(prop=n/sum(n))

resume |> 
  count(gender_fac) |> 
  mutate(prop=n/sum(n))

resume |> 
  group_by(received_callback_fac) |> 
  count(gender_fac) |> 
  mutate(prop=n/sum(n))

resume |>
  count(college_degree_fac) |>
  mutate(prop=n/sum(n))

resume |>
  group_by(received_callback_fac) |>
  count(college_degree_fac) |>
  mutate(prop=n/sum(n))

resume |>
  count(resume_quality_fac) |>
  mutate(prop=n/sum(n))

resume |>
  group_by(received_callback_fac) |>
  count(resume_quality_fac) |>
  mutate(prop=n/sum(n))

resume |>
  summarise(mean(years_experience), sd(years_experience))

resume |>
  group_by(received_callback_fac) |>
  summarise(mean(years_experience), sd(years_experience))
```

+---------------------------------+-----------+-------------------+--------------------------+
| Variable                        | Overall   | Received Callback | Did not receive callback |
|                                 |           |                   |                          |
|                                 | 4870      | 392 (8)           | 4478 (92)                |
+=================================+===========+===================+==========================+
| Race                            |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Black - N (%)                   | 2435 (50) | 157 (40)          | 2278 (51)                |
|                                 |           |                   |                          |
| White - N (%)                   | 2435 (50) | 235 (60)          | 2200 (49)                |
+---------------------------------+-----------+-------------------+--------------------------+
| Gender                          |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Female - N (%)                  | 3746 (77) | 309 (79)          | 3437 (77)                |
|                                 |           |                   |                          |
| Male - N (%)                    | 1124 (23) | 83 (21)           | 1041 (23)                |
+---------------------------------+-----------+-------------------+--------------------------+
| Has college degree - N (%)      | 3504 (72) | 277 (71)          | 3227 (72)                |
+---------------------------------+-----------+-------------------+--------------------------+
| Resume Quality                  |           |                   |                          |
+---------------------------------+-----------+-------------------+--------------------------+
| Low - N (%)                     | 2424 (50) | 178 (45)          | 2246 (50)                |
|                                 |           |                   |                          |
| High - N (%)                    | 2446 (50) | 214 (55)          | 2232 (50)                |
+---------------------------------+-----------+-------------------+--------------------------+
| Years of experience - mean (SD) | 7.8 (5)   | 8.9 (5.5)         | 7.8 (5)                  |
+---------------------------------+-----------+-------------------+--------------------------+

Takeaways: Most variables do not seem to have a big difference when it comes to receiving callbacks. The exception to this is race. While White applicants make up 50% of the experimental sample, they receive 60% of the callbacks. Callback rates align with the overall sample for gender. Interestingly, resume quality seems to have little effect, with low quality resumes receiving 45% of the callbacks.

## 2

a\.

```{r}
resume_mod1 <- glm(received_callback_fac ~
                     race_fac + gender_fac + college_degree_fac +
                     resume_quality_fac + years_experience,
                   data=resume,
                   family="binomial")

summary(resume_mod1)

exp(coefficients(resume_mod1))
```

b\.

```{r}
exp(confint(resume_mod1))
```

c\.

Gender, having a college degree, and resume quality are not statistically significantly associated with receiving a resume callback (p=0.4, 0.8, and 0.2, respectively). The odds of receiving a callback are 1.6 times higher for White applicants compared to Black applicants, and this is statistically significant (p\<0.001, 95% CI: \[1.3, 1.9\]). Years of experience is also significantly associated with receiving a callback; per additional year of experience, the odds of receiving a callback increase by 4% (p\<0.001, 95% CI: \[1.01,1.06\]).

## 3

a\.

```{r}
resume_mod2 <- glm(received_callback_fac ~
                     race_fac*gender_fac + college_degree_fac +
                     resume_quality_fac + years_experience,
                   data=resume,
                   family="binomial")

summary(resume_mod2)

exp(coefficients(resume_mod2))

exp(confint(resume_mod2))
```

The reference group is Black females. Let $x_1=1$ for race=White, 0=Black; $x_2=1$ for gender=male, 0=female.

-   For White males, $x_1=1$ and $x_2=1$, so the log odds of receiving a callback = -2.87+0.44-0.11+0.002=-2.5, so the odds of receiving a callback are $exp(-2.5)=0.08$.

-   For White females, $x_1=1$ and $x_2=0$, so the log odds of receiving a callback=-2.87+0.44, so the odds of receiving a callback are $exp(-2.4)=0.09$.

-   For Black males, $x_1=0$ and $x_2=1$, so the log odds of receiving a callback = -2.87-0.11, so the odds of receiving a callback are $exp(-2.98)=0.05$

-   For Black females, $x_1=0$ and $x_2=0$, so the log odds of receiving a callback = -2.87 (the intercept!), so the odds of receiving a callback are $exp(-2.9)=0.06$

So, White females have the highest odds of receiving a callback.

b\.

```{r}
resumemod_preds <- ifelse(predict(resume_mod2, type="response")>0.5,1,0)
resumemod_preds_fac <- factor(resumemod_preds,
                            levels=c(0,1),
                            labels=c("No", "Yes"))
confusionMatrix(table(resumemod_preds_fac, resume$received_callback_fac),
                positive="Yes",
                mode="everything")
```

Using the 0.5 threshold, no observations are predicted to get a callback. Although the accuracy is high (0.92), the kappa is 0 due to the imbalance in the data (only 8% receive a callback).

c\.

```{r}
roc(resume$received_callback_fac, predict(resume_mod2, type="response"), 
    print.thres="best", 
    print.auc=T,
    legacy.axes=T,
    plot=T)
```

```{r}
resumemod_preds_bestthres <- ifelse(predict(resume_mod2, type="response")>0.087,1,0)
resumemod_preds_fac_bestthres <- factor(resumemod_preds_bestthres,
                            levels=c(0,1),
                            labels=c("No", "Yes"))
confusionMatrix(table(resumemod_preds_fac_bestthres,
                      resume$received_callback_fac),
                positive="Yes",
                mode="everything")
```

We now have observations that were predicted to receive a callback. The accuracy decreased, but the kappa increased. The sensitivity and specificity are fairly low at 0.49 and 0.67, respectively. Given that the AUC is 0.58, this may not be a great model for prediction, but we can still trust our inferences!

## 4

Based on both models (with and without interaction), race and years of experience are statistically significantly associated with receiving a callback. Gender is not significant, and the interaction between race and gender is not significant. Looking at odds of receiving a callback for each race/gender combination, we see that White females and males have odds of .09 and .08, respectively, and Black females and males have odds of 0.06 and 0.05, respectively. The strength of the model in terms of predictive accuracy is not very good; this can be difficult with a fairly rare outcome like we have in this case. Looking at the data available, it could be interesting to look at different kinds of jobs and whether or not race and gender interact with job types and receiving a callback. For example, it could be interesting to see if men are more likely to receive a callback for titles that imply leadership, like managerial roles.
