---
title: "IDS 702 HW 5"
author: "KEY"
format: pdf
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute. You must show your work for all problems, and you must provide a written answer for all problems; it is insufficient to just show the code.

**Add your name in the Author section in the header**

## Load data

```{r}
library(MASS)
library(tidyverse)
library(glmmTMB)
## NOTE: Run install.packages("glmmTMB", type = "source") IN CONSOLE

disaster_deaths <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/disaster_deaths_cleaned.csv")
```

## 1

a\.

```{r}
ggplot(disaster_deaths, aes(x=Year, y=Deaths, col=Region))+
  geom_point()
```

We see an outlier that stands out from the rest of the data, and we see that many of the points are very small or zero. It is difficult to see a clear trend over time.

b\.

```{r}
disaster_deaths |>
  mutate(event = ifelse(Deaths>0,1,0)) |>
  count(event) |>
  mutate(prop=n/sum(n))
```

8109 (71%) of the rows are 0, and 3276 (29%) of the rows are nonzero.

## 2

a\.

```{r}
disaster_pois_mod <- glm(Deaths ~ Region*Year,
                         data=disaster_deaths,
                         family="poisson")

summary(d